---
title: "Upper Gila Watershed Bayesian Network Implementation"
author:"Lauren Victoria Jaramillo, UNM (alwayslo@unm.edu)"
date: "4/2/2019"
update: "4/4/2019"
output: html_notebook
---

This R notebook is used to used to implement the constructed Bayesian network for riparian recruitment for various scenarios (natural, scenario 1, and scenario 2). These scenarios corresponde to the CUFA.

This first code chunck clears the environment, installs all the necessary packages, and returns the current working directory to ensure filespaths will execute properly.This code chunck loads and attaches the add-on packages *bnlearn*,*readr*,and *Rgraphviz*.*Rgraphviz* requires additional packages *BioGenerics* and graph to run  It also clears the the global environment of any previous data.
```{r}
#rm(list=ls())

library("bnlearn")
library("readr")
library("Rgraphviz")
library("xts")
library("prob")
library("TTR")
library("bigmemory")
library("zoo")
library("raster")
library("sp")
library("rgdal")
library("rasterVis")
library("ggplot2")
library("rgdal")
library("dplyr")
library("methods")
require("BiocGenerics")
require("graph")

getwd()
```

###Import SRH-2D model results###
This code chunck calls the txt file that includes the inundation stats for each inundated cell in the hydrodynamic model mesh. This calls the data for one of 5 sites.

```{r}
q_cells <- read.table("data/site1_1000_to_4000.txt",header = FALSE)
data.frame(q_cells)
q_cells <- setNames(q_cells, c("cell","flow"))

q_cells$bin <-    ifelse(q_cells$flow <= q_bin[1],"bin_1",
                    ifelse(q_bin[1]< q_cells$flow  & q_cells$flow <=q_bin[2],"bin_2",
                    ifelse(q_bin[2]< q_cells$flow  & q_cells$flow <=q_bin[3],"bin_3",
                    ifelse(q_bin[3]< q_cells$flow  & q_cells$flow <=q_bin[4],"bin_4",
                    ifelse(q_bin[4]< q_cells$flow  & q_cells$flow <=q_bin[5],"bin_5",
                    ifelse(q_bin[5]< q_cells$flow  & q_cells$flow <=q_bin[6],"bin_6",
                    ifelse(q_bin[6]< q_cells$flow  & q_cells$flow <=q_bin[7],"bin_7",
                    ifelse(q_cells$flow >q_bin[7],"bin_8","blah"))))))))

```


##Natural Scenario Evidence##
```{r}
Evidence <- c(hydro[1:nrow(recessionrate_14d),],recessionrate_14d[,2:4])
Evidence <-data.frame(Evidence)
Evidence$TIMING <- NA
Evidence$INUN <- NA
Evidence$RECESSION_RATE<- NA
Evidence <- setNames(Evidence, c("date","q_natural", "q_scenario1","q_scenario2","rr_natural", "rr_scenario1","rr_scenario2", "TIMING","INUN","RECESSION_RATE"))

Evidence$date<-as.Date(Evidence$date, format="%m/%d/%Y")

Evidence$TIMING <-  ifelse(months(Evidence$date) %in% month.name[4:5],"apr_may",
                    ifelse(months(Evidence$date) %in% month.name[6:7],"jun_jul",
                    ifelse(months(Evidence$date) %in% month.name[8:9],"aug_sep","blah")))

Evidence$INUN <-    ifelse(Evidence$q_natural<= q_bin[1],"bin_1",
                    ifelse(q_bin[1]< Evidence$q_natural & Evidence$q_natural<=q_bin[2],"bin_2",
                    ifelse(q_bin[2]< Evidence$q_natural & Evidence$q_natural<=q_bin[3],"bin_3",
                    ifelse(q_bin[3]< Evidence$q_natural & Evidence$q_natural<=q_bin[4],"bin_4",
                    ifelse(q_bin[4]< Evidence$q_natural & Evidence$q_natural<=q_bin[5],"bin_5",
                    ifelse(q_bin[5]< Evidence$q_natural & Evidence$q_natural<=q_bin[6],"bin_6",
                    ifelse(q_bin[6]< Evidence$q_natural & Evidence$q_natural<=q_bin[7],"bin_7",
                    ifelse(Evidence$q_natural>q_bin[7],"bin_8","blah"))))))))


Evidence$RECESSION_RATE <- ifelse(Evidence$rr_natural <= recessionrate_bin[1],"bin_1",
                           ifelse(recessionrate_bin[1]< Evidence$rr_natural & Evidence$rr_natural <=recessionrate_bin[2],"bin_2",
                           ifelse(recessionrate_bin[2]< Evidence$rr_natural & Evidence$rr_natural <=recessionrate_bin[3],"bin_3",
                           ifelse(recessionrate_bin[3]< Evidence$rr_natural & Evidence$rr_natural <=recessionrate_bin[4],"bin_4",
                           ifelse(Evidence$rr_natural > recessionrate_bin[4],"bin_5","blah")))))  

Evidence <- Evidence[-1:-7]
```

##Scenario 1 Evidence##
```{r}
Evidence <- c(hydro[1:nrow(recessionrate_14d),],recessionrate_14d[,2:4])
Evidence <-data.frame(Evidence)
Evidence$TIMING <- NA
Evidence$INUN <- NA
Evidence$RECESSION_RATE<- NA
Evidence <- setNames(Evidence, c("date","q_natural", "q_scenario1","q_scenario2","rr_natural", "rr_scenario1","rr_scenario2", "TIMING","INUN","RECESSION_RATE"))

Evidence$date<-as.Date(Evidence$date, format="%m/%d/%Y")

Evidence$TIMING <-  ifelse(months(Evidence$date) %in% month.name[4:5],"apr_may",
                    ifelse(months(Evidence$date) %in% month.name[6:7],"jun_jul",
                    ifelse(months(Evidence$date) %in% month.name[8:9],"aug_sep","blah")))

Evidence$INUN <-    ifelse(Evidence$q_scenario1<= q_bin[1],"bin_1",
                    ifelse(q_bin[1]< Evidence$q_scenario1 & Evidence$q_scenario1<=q_bin[2],"bin_2",
                    ifelse(q_bin[2]< Evidence$q_scenario1 & Evidence$q_scenario1<=q_bin[3],"bin_3",
                    ifelse(q_bin[3]< Evidence$q_scenario1 & Evidence$q_scenario1<=q_bin[4],"bin_4",
                    ifelse(q_bin[4]< Evidence$q_scenario1 & Evidence$q_scenario1<=q_bin[5],"bin_5",
                    ifelse(q_bin[5]< Evidence$q_scenario1 & Evidence$q_scenario1<=q_bin[6],"bin_6",
                    ifelse(q_bin[6]< Evidence$q_scenario1 & Evidence$q_scenario1<=q_bin[7],"bin_7",
                    ifelse(Evidence$q_scenario1>q_bin[7],"bin_8","blah"))))))))


Evidence$RECESSION_RATE <- ifelse(Evidence$rr_natural <= recessionrate_bin[1],"bin_1",
                           ifelse(recessionrate_bin[1]< Evidence$rr_natural & Evidence$rr_natural <=recessionrate_bin[2],"bin_2",
                           ifelse(recessionrate_bin[2]< Evidence$rr_natural & Evidence$rr_natural <=recessionrate_bin[3],"bin_3",
                           ifelse(recessionrate_bin[3]< Evidence$rr_natural & Evidence$rr_natural <=recessionrate_bin[4],"bin_4",
                           ifelse(Evidence$rr_natural > recessionrate_bin[4],"bin_5","blah")))))  

Evidence <- Evidence[-1:-7]
```

##Scenario 2 Evidence##
```{r}
Evidence <- c(hydro[1:nrow(recessionrate_14d),],recessionrate_14d[,2:4])
Evidence <-data.frame(Evidence)
Evidence$TIMING <- NA
Evidence$INUN <- NA
Evidence$RECESSION_RATE<- NA
Evidence <- setNames(Evidence, c("date","q_natural", "q_scenario1","q_scenario2","rr_natural", "rr_scenario1","rr_scenario2", "TIMING","INUN","RECESSION_RATE"))

Evidence$date<-as.Date(Evidence$date, format="%m/%d/%Y")

Evidence$TIMING <-  ifelse(months(Evidence$date) %in% month.name[4:5],"apr_may",
                    ifelse(months(Evidence$date) %in% month.name[6:7],"jun_jul",
                    ifelse(months(Evidence$date) %in% month.name[8:9],"aug_sep","blah")))

Evidence$INUN <-    ifelse(Evidence$q_scenario2<= q_bin[1],"bin_1",
                    ifelse(q_bin[1]< Evidence$q_scenario2 & Evidence$q_scenario2<=q_bin[2],"bin_2",
                    ifelse(q_bin[2]< Evidence$q_scenario2 & Evidence$q_scenario2<=q_bin[3],"bin_3",
                    ifelse(q_bin[3]< Evidence$q_scenario2 & Evidence$q_scenario2<=q_bin[4],"bin_4",
                    ifelse(q_bin[4]< Evidence$q_scenario2 & Evidence$q_scenario2<=q_bin[5],"bin_5",
                    ifelse(q_bin[5]< Evidence$q_scenario2 & Evidence$q_scenario2<=q_bin[6],"bin_6",
                    ifelse(q_bin[6]< Evidence$q_scenario2 & Evidence$q_scenario2<=q_bin[7],"bin_7",
                    ifelse(Evidence$q_scenario2>q_bin[7],"bin_8","blah"))))))))


Evidence$RECESSION_RATE <- ifelse(Evidence$rr_natural <= recessionrate_bin[1],"bin_1",
                           ifelse(recessionrate_bin[1]< Evidence$rr_natural & Evidence$rr_natural <=recessionrate_bin[2],"bin_2",
                           ifelse(recessionrate_bin[2]< Evidence$rr_natural & Evidence$rr_natural <=recessionrate_bin[3],"bin_3",
                           ifelse(recessionrate_bin[3]< Evidence$rr_natural & Evidence$rr_natural <=recessionrate_bin[4],"bin_4",
                           ifelse(Evidence$rr_natural > recessionrate_bin[4],"bin_5","blah")))))  

Evidence <- Evidence[-1:-7]
```

```{r}
Evidence_q1 <- Evidence[Evidence[2] =="bin_1" & Evidence[1]!="blah",]
Evidence_q2 <- Evidence[Evidence[2] =="bin_2" & Evidence[1]!="blah",]
Evidence_q3 <- Evidence[Evidence[2] =="bin_3" & Evidence[1]!="blah",]
Evidence_q4 <- Evidence[Evidence[2] =="bin_4" & Evidence[1]!="blah",]
Evidence_q5 <- Evidence[Evidence[2] =="bin_5" & Evidence[1]!="blah",]
Evidence_q6 <- Evidence[Evidence[2] =="bin_6" & Evidence[1]!="blah",]
Evidence_q7 <- Evidence[Evidence[2] =="bin_7" & Evidence[1]!="blah",]
Evidence_q8 <- Evidence[Evidence[2] =="bin_8" & Evidence[1]!="blah",]
```

#### Calculate recruitment probability by instantiating the BN with evidence ####
```{r}
Recruit_q1<-data.frame()
for (i in 1:(nrow(Evidence_q1))) 
{Recruit_q1 <-rbind(Recruit_q1, cpquery(bnfit,event = (RECRUIT =="YES"),
                                        evidence = list(TIMING=Evidence_q1[i,1],
                                                        INUN=Evidence_q1[i,2],
                                                        RECESSION_RATE = Evidence_q1[i,3]), 
                                        method = "lw"))}

Recruit_q2<-data.frame()
for (i in 1:(nrow(Evidence_q2))) 
{Recruit_q2 <-rbind(Recruit_q2, cpquery(bnfit,event = (RECRUIT =="YES"),
                                        evidence = list(TIMING=Evidence_q2[i,1],
                                                        INUN=Evidence_q2[i,2],
                                                        RECESSION_RATE = Evidence_q2[i,3]), 
                                        method = "lw"))}

Recruit_q3<-data.frame()
for (i in 1:(nrow(Evidence_q3))) 
{Recruit_q3 <-rbind(Recruit_q3, cpquery(bnfit,event = (RECRUIT =="YES"),
                                        evidence = list(TIMING=Evidence_q3[i,1],
                                                        INUN=Evidence_q3[i,2],
                                                        RECESSION_RATE = Evidence_q3[i,3]), 
                                        method = "lw"))}

Recruit_q4<-data.frame()
for (i in 1:(nrow(Evidence_q4))) 
{Recruit_q4 <-rbind(Recruit_q4, cpquery(bnfit,event = (RECRUIT =="YES"),
                                        evidence = list(TIMING=Evidence_q4[i,1],
                                                        INUN=Evidence_q4[i,2],
                                                        RECESSION_RATE = Evidence_q4[i,3]), 
                                        method = "lw"))}

Recruit_q5<-data.frame()
for (i in 1:(nrow(Evidence_q5))) 
{Recruit_q5 <-rbind(Recruit_q5, cpquery(bnfit,event = (RECRUIT =="YES"),
                                        evidence = list(TIMING=Evidence_q5[i,1],
                                                        INUN=Evidence_q5[i,2],
                                                        RECESSION_RATE = Evidence_q5[i,3]), 
                                        method = "lw"))}

Recruit_q6<-data.frame()
for (i in 1:(nrow(Evidence_q6))) 
{Recruit_q6 <-rbind(Recruit_q6, cpquery(bnfit,event = (RECRUIT =="YES"),
                                        evidence = list(TIMING=Evidence_q6[i,1],
                                                        INUN=Evidence_q6[i,2],
                                                        RECESSION_RATE = Evidence_q6[i,3]), 
                                        method = "lw"))}

Recruit_q7<-data.frame()
for (i in 1:(nrow(Evidence_q7))) 
{Recruit_q7 <-rbind(Recruit_q7, cpquery(bnfit,event = (RECRUIT =="YES"),
                                        evidence = list(TIMING=Evidence_q7[i,1],
                                                        INUN=Evidence_q7[i,2],
                                                        RECESSION_RATE = Evidence_q7[i,3]), 
                                        method = "lw"))}

Recruit_q8<-data.frame()
for (i in 1:(nrow(Evidence_q8))) 
{Recruit_q8 <-rbind(Recruit_q8, cpquery(bnfit,event = (RECRUIT =="YES"),
                                        evidence = list(TIMING=Evidence_q8[i,1],
                                                        INUN=Evidence_q8[i,2],
                                                        RECESSION_RATE = Evidence_q8[i,3]), 
                                        method = "lw"))}
```

Two output files,(1) The recruitment probability for each q_bin and (2) the recruitment probability for each cell, are generated, written, and saved in the output folder. Result/output files can be re-written to specify which scenario is being written to .csv for clarity (e.g.q_cells_natural_recruit.csv ).The file path can also be changed as needed.
```{r}
Recruit_prob <- c(mean(Recruit_q1[,1]),mean(Recruit_q2[,1]),mean(Recruit_q3[,1]),mean(Recruit_q4[,1]),mean(Recruit_q5[,1]),mean(Recruit_q6[,1]),mean(Recruit_q7[,1]),mean(Recruit_q8[,1]))
Recruit_prob<-data.frame(Recruit_prob)
dimnames(Recruit_prob) = list("bin"=c("bin_1", "bin_2", "bin_3", "bin_4", "bin_5","bin_6","bin_7", "bin_8"),"Recruit")

q_cells$recruit <-    ifelse(q_cells$bin =="bin_1",Recruit_prob[1,],
                    ifelse(q_cells$bin =="bin_2",Recruit_prob[2,],
                    ifelse(q_cells$bin =="bin_3",Recruit_prob[3,],
                    ifelse(q_cells$bin =="bin_4",Recruit_prob[4,],
                    ifelse(q_cells$bin =="bin_5",Recruit_prob[5,],
                    ifelse(q_cells$bin =="bin_6",Recruit_prob[6,],
                    ifelse(q_cells$bin =="bin_7",Recruit_prob[7,],
                    Recruit_prob[8,])))))))

write.csv(Recruit_prob,file = "/Volumes/LoBear/BN/Gila/bn_master/output/old_natural2_recruit_prob.csv")
write.csv(q_cells,file = "/Volumes/LoBear/BN/Gila/bn_master/output/old_natural2_q_cells_recruit.csv",row.names= FALSE)
```

####R Notebooks####
Execute code chunks by clicking the *Run* button within the chunk or by placing the cursor inside it and pressing *Cmd+Shift+Enter*. Add a new chunk by clicking the *Insert Chunk* button on the toolbar or by pressing *Cmd+Option+I*. Whenthe notebook is saved, an HTML file containing the code and output will be saved alongside it (click the *Preview* button or press *Cmd+Shift+K* to preview the HTML file).

####References ####
Gentleman R, Whalen E, Huber W and Falcon S (2017). graph: graph: A package to handle graph data structures. R package version 1.54.0.

Hansen KD, Gentry J, Long L, Gentleman R, Falcon S, Hahne F and Sarkar D (2017). Rgraphviz: Provides plotting capabilities for R graph objects. R package version 2.20.0.

Huber, W., Carey, J. V, Gentleman, R., Anders, S., Carlson, M., Carvalho, S. B, Bravo, C. H, Davis, S., Gatto, L., Girke, T., Gottardo, R., Hahne, F., Hansen, D. K, Irizarry, A. R, Lawrence, M., Love, I. M, MacDonald, J., Obenchain, V., Ole's, K. A, Pag'es, H., Reyes, A., Shannon, P., Smyth, K. G, Tenenbaum, D., Waldron, L., Morgan and M. (2015). “Orchestrating high-throughput genomic analysis with Bioconductor.” Nature Methods, 12(2), pp. 115–121. http://www.nature.com/nmeth/journal/v12/n2/full/nmeth.3252.html.

Koller D, Friedman N (2009). Probabilistic Graphical Models: Principles and Techniques. MIT Press.

Pearl J (2009). Causality: Models, Reasoning and Inference. Cambridge University Press, 2nd edition.

Nagarajan, Radhakrishnan, Marco Scutari, and Sophie Lèbre. Bayesian Networks in R. New York, NY: Springer New York, 2013. http://link.springer.com/10.1007/978-1-4614-6446-4. 
http://www.bnlearn.com

Scutari, Marco (2018). bnlearn: Bayesian Network Structure Learning, Parameter Learning and Inference. R package version 4.4 https://cran.r-project.org/web/packages/bnlearn/

https://stackoverflow.com/questions/40364458/r-some-conditional-probability-distributions-of-node-do-not-sum-to-one

example - https://stackoverflow.com/questions/34127885/bnlearn-wrong-dimensions-for-node

cpquery - http://www.bnlearn.com/documentation/man/cpquery.html

soft/hard evidence - https://pdfs.semanticscholar.org/13f5/795c9d4075d4468131fe9405bae5ec3e3857.pdf

