---
title: "Upper Gila Watershed Bayesian Network Implementation"
author:"Lauren Victoria Jaramillo, UNM (alwayslo@unm.edu)"
date: "4/2/2019"
update: "4/4/2019"
output: html_notebook
---

This R notebook is used to used to implement the constructed Bayesian network for riparian recruitment for various scenarios (natural, scenario 1, and scenario 2). These scenarios corresponde to the CUFA.

This first code chunck clears the environment, installs all the necessary packages, and returns the current working directory to ensure filespaths will execute properly.This code chunck loads and attaches the add-on packages *bnlearn*,*readr*,and *Rgraphviz*.*Rgraphviz* requires additional packages *BioGenerics* and graph to run  It also clears the the global environment of any previous data.
```{r}
#rm(list=ls())

library("bnlearn")
library("readr")
library("Rgraphviz")
library("xts")
library("prob")
library("TTR")
library("bigmemory")
library("zoo")
library("raster")
library("sp")
library("rgdal")
library("rasterVis")
library("ggplot2")
library("rgdal")
library("dplyr")
library("methods")
require("BiocGenerics")
require("graph")

getwd()
```

###Import SRH-2D model results###
This code chunck calls the txt file that includes the inundation stats for each inundated cell in the hydrodynamic model mesh. This calls the data for one of 5 sites.

The 
```{r}
q_cells <- read.table("data/site1_1000_to_4000.txt",header = FALSE)
data.frame(q_cells)
q_cells <- setNames(q_cells, c("cell","flow"))

q_cells$bin <-    ifelse(q_cells$flow <= q_bin[1],"bin_1",
                    ifelse(q_bin[1]< q_cells$flow  & q_cells$flow <=q_bin[2],"bin_2",
                    ifelse(q_bin[2]< q_cells$flow  & q_cells$flow <=q_bin[3],"bin_3",
                    ifelse(q_bin[3]< q_cells$flow  & q_cells$flow <=q_bin[4],"bin_4",
                    ifelse(q_bin[4]< q_cells$flow  & q_cells$flow <=q_bin[5],"bin_5",
                    ifelse(q_bin[5]< q_cells$flow  & q_cells$flow <=q_bin[6],"bin_6",
                    ifelse(q_bin[6]< q_cells$flow  & q_cells$flow <=q_bin[7],"bin_7",
                    ifelse(q_cells$flow >q_bin[7],"bin_8","blah"))))))))

```


##Natural Scenario Evidence##
```{r}
Evidence <- c(hydro[1:nrow(recessionrate_14d),],recessionrate_14d[,2:4])
Evidence <-data.frame(Evidence)
Evidence$TIMING <- NA
Evidence$INUN <- NA
Evidence$RECESSION_RATE<- NA
Evidence <- setNames(Evidence, c("date","q_natural", "q_scenario1","q_scenario2","rr_natural", "rr_scenario1","rr_scenario2", "TIMING","INUN","RECESSION_RATE"))

Evidence$date<-as.Date(Evidence$date, format="%m/%d/%Y")

Evidence$TIMING <-  ifelse(months(Evidence$date) %in% month.name[4:5],"apr_may",
                    ifelse(months(Evidence$date) %in% month.name[6:7],"jun_jul",
                    ifelse(months(Evidence$date) %in% month.name[8:9],"aug_sep","blah")))

Evidence$INUN <-    ifelse(Evidence$q_natural<= q_bin[1],"bin_1",
                    ifelse(q_bin[1]< Evidence$q_natural & Evidence$q_natural<=q_bin[2],"bin_2",
                    ifelse(q_bin[2]< Evidence$q_natural & Evidence$q_natural<=q_bin[3],"bin_3",
                    ifelse(q_bin[3]< Evidence$q_natural & Evidence$q_natural<=q_bin[4],"bin_4",
                    ifelse(q_bin[4]< Evidence$q_natural & Evidence$q_natural<=q_bin[5],"bin_5",
                    ifelse(q_bin[5]< Evidence$q_natural & Evidence$q_natural<=q_bin[6],"bin_6",
                    ifelse(q_bin[6]< Evidence$q_natural & Evidence$q_natural<=q_bin[7],"bin_7",
                    ifelse(Evidence$q_natural>q_bin[7],"bin_8","blah"))))))))


Evidence$RECESSION_RATE <- ifelse(Evidence$rr_natural <= recessionrate_bin[1],"bin_1",
                           ifelse(recessionrate_bin[1]< Evidence$rr_natural & Evidence$rr_natural <=recessionrate_bin[2],"bin_2",
                           ifelse(recessionrate_bin[2]< Evidence$rr_natural & Evidence$rr_natural <=recessionrate_bin[3],"bin_3",
                           ifelse(recessionrate_bin[3]< Evidence$rr_natural & Evidence$rr_natural <=recessionrate_bin[4],"bin_4",
                           ifelse(Evidence$rr_natural > recessionrate_bin[4],"bin_5","blah")))))  

Evidence <- Evidence[-1:-7]
```

##Scenario 1 Evidence##
```{r}
Evidence <- c(hydro[1:nrow(recessionrate_14d),],recessionrate_14d[,2:4])
Evidence <-data.frame(Evidence)
Evidence$TIMING <- NA
Evidence$INUN <- NA
Evidence$RECESSION_RATE<- NA
Evidence <- setNames(Evidence, c("date","q_natural", "q_scenario1","q_scenario2","rr_natural", "rr_scenario1","rr_scenario2", "TIMING","INUN","RECESSION_RATE"))

Evidence$date<-as.Date(Evidence$date, format="%m/%d/%Y")

Evidence$TIMING <-  ifelse(months(Evidence$date) %in% month.name[4:5],"apr_may",
                    ifelse(months(Evidence$date) %in% month.name[6:7],"jun_jul",
                    ifelse(months(Evidence$date) %in% month.name[8:9],"aug_sep","blah")))

Evidence$INUN <-    ifelse(Evidence$q_scenario1<= q_bin[1],"bin_1",
                    ifelse(q_bin[1]< Evidence$q_scenario1 & Evidence$q_scenario1<=q_bin[2],"bin_2",
                    ifelse(q_bin[2]< Evidence$q_scenario1 & Evidence$q_scenario1<=q_bin[3],"bin_3",
                    ifelse(q_bin[3]< Evidence$q_scenario1 & Evidence$q_scenario1<=q_bin[4],"bin_4",
                    ifelse(q_bin[4]< Evidence$q_scenario1 & Evidence$q_scenario1<=q_bin[5],"bin_5",
                    ifelse(q_bin[5]< Evidence$q_scenario1 & Evidence$q_scenario1<=q_bin[6],"bin_6",
                    ifelse(q_bin[6]< Evidence$q_scenario1 & Evidence$q_scenario1<=q_bin[7],"bin_7",
                    ifelse(Evidence$q_scenario1>q_bin[7],"bin_8","blah"))))))))


Evidence$RECESSION_RATE <- ifelse(Evidence$rr_natural <= recessionrate_bin[1],"bin_1",
                           ifelse(recessionrate_bin[1]< Evidence$rr_natural & Evidence$rr_natural <=recessionrate_bin[2],"bin_2",
                           ifelse(recessionrate_bin[2]< Evidence$rr_natural & Evidence$rr_natural <=recessionrate_bin[3],"bin_3",
                           ifelse(recessionrate_bin[3]< Evidence$rr_natural & Evidence$rr_natural <=recessionrate_bin[4],"bin_4",
                           ifelse(Evidence$rr_natural > recessionrate_bin[4],"bin_5","blah")))))  

Evidence <- Evidence[-1:-7]
```

##Scenario 2 Evidence##
```{r}
Evidence <- c(hydro[1:nrow(recessionrate_14d),],recessionrate_14d[,2:4])
Evidence <-data.frame(Evidence)
Evidence$TIMING <- NA
Evidence$INUN <- NA
Evidence$RECESSION_RATE<- NA
Evidence <- setNames(Evidence, c("date","q_natural", "q_scenario1","q_scenario2","rr_natural", "rr_scenario1","rr_scenario2", "TIMING","INUN","RECESSION_RATE"))

Evidence$date<-as.Date(Evidence$date, format="%m/%d/%Y")

Evidence$TIMING <-  ifelse(months(Evidence$date) %in% month.name[4:5],"apr_may",
                    ifelse(months(Evidence$date) %in% month.name[6:7],"jun_jul",
                    ifelse(months(Evidence$date) %in% month.name[8:9],"aug_sep","blah")))

Evidence$INUN <-    ifelse(Evidence$q_scenario2<= q_bin[1],"bin_1",
                    ifelse(q_bin[1]< Evidence$q_scenario2 & Evidence$q_scenario2<=q_bin[2],"bin_2",
                    ifelse(q_bin[2]< Evidence$q_scenario2 & Evidence$q_scenario2<=q_bin[3],"bin_3",
                    ifelse(q_bin[3]< Evidence$q_scenario2 & Evidence$q_scenario2<=q_bin[4],"bin_4",
                    ifelse(q_bin[4]< Evidence$q_scenario2 & Evidence$q_scenario2<=q_bin[5],"bin_5",
                    ifelse(q_bin[5]< Evidence$q_scenario2 & Evidence$q_scenario2<=q_bin[6],"bin_6",
                    ifelse(q_bin[6]< Evidence$q_scenario2 & Evidence$q_scenario2<=q_bin[7],"bin_7",
                    ifelse(Evidence$q_scenario2>q_bin[7],"bin_8","blah"))))))))


Evidence$RECESSION_RATE <- ifelse(Evidence$rr_natural <= recessionrate_bin[1],"bin_1",
                           ifelse(recessionrate_bin[1]< Evidence$rr_natural & Evidence$rr_natural <=recessionrate_bin[2],"bin_2",
                           ifelse(recessionrate_bin[2]< Evidence$rr_natural & Evidence$rr_natural <=recessionrate_bin[3],"bin_3",
                           ifelse(recessionrate_bin[3]< Evidence$rr_natural & Evidence$rr_natural <=recessionrate_bin[4],"bin_4",
                           ifelse(Evidence$rr_natural > recessionrate_bin[4],"bin_5","blah")))))  

Evidence <- Evidence[-1:-7]
```

```{r}
Evidence_q1 <- Evidence[Evidence[2] =="bin_1" & Evidence[1]!="blah",]
Evidence_q2 <- Evidence[Evidence[2] =="bin_2" & Evidence[1]!="blah",]
Evidence_q3 <- Evidence[Evidence[2] =="bin_3" & Evidence[1]!="blah",]
Evidence_q4 <- Evidence[Evidence[2] =="bin_4" & Evidence[1]!="blah",]
Evidence_q5 <- Evidence[Evidence[2] =="bin_5" & Evidence[1]!="blah",]
Evidence_q6 <- Evidence[Evidence[2] =="bin_6" & Evidence[1]!="blah",]
Evidence_q7 <- Evidence[Evidence[2] =="bin_7" & Evidence[1]!="blah",]
Evidence_q8 <- Evidence[Evidence[2] =="bin_8" & Evidence[1]!="blah",]
```

#### Calculate recruitment probability by instantiating the BN with evidence ####
```{r}
Recruit_q1<-data.frame()
for (i in 1:(nrow(Evidence_q1))) 
{Recruit_q1 <-rbind(Recruit_q1, cpquery(bnfit,event = (RECRUIT =="YES"),
                                        evidence = list(TIMING=Evidence_q1[i,1],
                                                        INUN=Evidence_q1[i,2],
                                                        RECESSION_RATE = Evidence_q1[i,3]), 
                                        method = "lw"))}

Recruit_q2<-data.frame()
for (i in 1:(nrow(Evidence_q2))) 
{Recruit_q2 <-rbind(Recruit_q2, cpquery(bnfit,event = (RECRUIT =="YES"),
                                        evidence = list(TIMING=Evidence_q2[i,1],
                                                        INUN=Evidence_q2[i,2],
                                                        RECESSION_RATE = Evidence_q2[i,3]), 
                                        method = "lw"))}

Recruit_q3<-data.frame()
for (i in 1:(nrow(Evidence_q3))) 
{Recruit_q3 <-rbind(Recruit_q3, cpquery(bnfit,event = (RECRUIT =="YES"),
                                        evidence = list(TIMING=Evidence_q3[i,1],
                                                        INUN=Evidence_q3[i,2],
                                                        RECESSION_RATE = Evidence_q3[i,3]), 
                                        method = "lw"))}

Recruit_q4<-data.frame()
for (i in 1:(nrow(Evidence_q4))) 
{Recruit_q4 <-rbind(Recruit_q4, cpquery(bnfit,event = (RECRUIT =="YES"),
                                        evidence = list(TIMING=Evidence_q4[i,1],
                                                        INUN=Evidence_q4[i,2],
                                                        RECESSION_RATE = Evidence_q4[i,3]), 
                                        method = "lw"))}

Recruit_q5<-data.frame()
for (i in 1:(nrow(Evidence_q5))) 
{Recruit_q5 <-rbind(Recruit_q5, cpquery(bnfit,event = (RECRUIT =="YES"),
                                        evidence = list(TIMING=Evidence_q5[i,1],
                                                        INUN=Evidence_q5[i,2],
                                                        RECESSION_RATE = Evidence_q5[i,3]), 
                                        method = "lw"))}

Recruit_q6<-data.frame()
for (i in 1:(nrow(Evidence_q6))) 
{Recruit_q6 <-rbind(Recruit_q6, cpquery(bnfit,event = (RECRUIT =="YES"),
                                        evidence = list(TIMING=Evidence_q6[i,1],
                                                        INUN=Evidence_q6[i,2],
                                                        RECESSION_RATE = Evidence_q6[i,3]), 
                                        method = "lw"))}

Recruit_q7<-data.frame()
for (i in 1:(nrow(Evidence_q7))) 
{Recruit_q7 <-rbind(Recruit_q7, cpquery(bnfit,event = (RECRUIT =="YES"),
                                        evidence = list(TIMING=Evidence_q7[i,1],
                                                        INUN=Evidence_q7[i,2],
                                                        RECESSION_RATE = Evidence_q7[i,3]), 
                                        method = "lw"))}

Recruit_q8<-data.frame()
for (i in 1:(nrow(Evidence_q8))) 
{Recruit_q8 <-rbind(Recruit_q8, cpquery(bnfit,event = (RECRUIT =="YES"),
                                        evidence = list(TIMING=Evidence_q8[i,1],
                                                        INUN=Evidence_q8[i,2],
                                                        RECESSION_RATE = Evidence_q8[i,3]), 
                                        method = "lw"))}
```


```{r}
Recruit_prob <- data.frame()
Recruit_prob <- cbind(Recruit_prob,mean(Recruit_q1[,1]),mean(Recruit_q2[,1]),mean(Recruit_q3[,1]),mean(Recruit_q4[,1]),mean(Recruit_q5[,1]),mean(Recruit_q6[,1]),mean(Recruit_q7[,1]),mean(Recruit_q8[,1])))

```

### Export File###
write.csv(E_q_all_prob_mn, file="output/site3/E_q_all_prob_mn.csv")
E_q_all_prob_mn2 <- read.csv("output/site3/E_q_all_prob_mn.csv")
colnames(E_q_all_prob_mn2) <- c("cell","mean_prob")
E_q_all_prob_mn2 <- E_q_all_prob_mn2[order(E_q_all_prob_mn2$cell),]
write.csv(E_q_all_prob_mn2, file="output/site3/E_q_all_prob_mn.csv", row.names=FALSE)

#### Calculate recruitment probability by instantiating the BN with evidence ####
system.time(E_q1_cell_probs <- lapply(E_evidence1, apply, 1, function(x) cpquery(riparian.fit1, (POTENTIAL=="Y"), evidence=as.list(x), method="lw")))
save(E_q1_cell_probs, file="output/site3/E_q1_cell_probs.Rdata")
system.time(E_q2_cell_probs <- lapply(E_evidence2, apply, 1, function(x) cpquery(riparian.fit2, (POTENTIAL=="Y"), evidence=as.list(x), method="lw")))
save(E_q2_cell_probs, file="output/site3/E_q2_cell_probs.Rdata")
system.time(E_q3_cell_probs <- lapply(E_evidence3, apply, 1, function(x) cpquery(riparian.fit3, (POTENTIAL=="Y"), evidence=as.list(x), method="lw")))
save(E_q3_cell_probs, file="output/site3/E_q3_cell_probs.Rdata")
system.time(E_q4_cell_probs <- lapply(E_evidence4, apply, 1, function(x) cpquery(riparian.fit4, (POTENTIAL=="Y"), evidence=as.list(x), method="lw")))
save(E_q4_cell_probs, file="output/site3/E_q4_cell_probs.Rdata")
system.time(E_q5_cell_probs <- lapply(E_evidence5, apply, 1, function(x) cpquery(riparian.fit5, (POTENTIAL=="Y"), evidence=as.list(x), method="lw")))
save(E_q5_cell_probs, file="output/site3/E_q5_cell_probs.Rdata")
system.time(E_q6_cell_probs <- lapply(E_evidence6, apply, 1, function(x) cpquery(riparian.fit6, (POTENTIAL=="Y"), evidence=as.list(x), method="lw")))
save(E_q6_cell_probs, file="output/site3/E_q6_cell_probs.Rdata")
system.time(E_q7_cell_probs <- lapply(E_evidence7, apply, 1, function(x) cpquery(riparian.fit7, (POTENTIAL=="Y"), evidence=as.list(x), method="lw")))
save(E_q7_cell_probs, file="output/site3/E_q7_cell_probs.Rdata")

E_q1_cell_prob_mn <- t(data.frame(lapply(E_q1_cell_probs, mean, na.rm=TRUE), check.names=FALSE))
E_q2_cell_prob_mn <- t(data.frame(lapply(E_q2_cell_probs, mean, na.rm=TRUE), check.names=FALSE))
E_q3_cell_prob_mn <- t(data.frame(lapply(E_q3_cell_probs, mean, na.rm=TRUE), check.names=FALSE))
E_q4_cell_prob_mn <- t(data.frame(lapply(E_q4_cell_probs, mean, na.rm=TRUE), check.names=FALSE))
E_q5_cell_prob_mn <- t(data.frame(lapply(E_q5_cell_probs, mean, na.rm=TRUE), check.names=FALSE))
E_q6_cell_prob_mn <- t(data.frame(lapply(E_q6_cell_probs, mean, na.rm=TRUE), check.names=FALSE))
E_q7_cell_prob_mn <- t(data.frame(lapply(E_q7_cell_probs, mean, na.rm=TRUE), check.names=FALSE))
if (exists("E_q7_cell_prob_mn")) E_q_all_prob_mn <- rbind(E_q1_cell_prob_mn, E_q2_cell_prob_mn, E_q3_cell_prob_mn, E_q4_cell_prob_mn, E_q5_cell_prob_mn, E_q6_cell_prob_mn, E_q7_cell_prob_mn)
if (!exists("E_q7_cell_prob_mn")) E_q_all_prob_mn <- rbind(E_q1_cell_prob_mn, E_q2_cell_prob_mn, E_q3_cell_prob_mn, E_q4_cell_prob_mn, E_q5_cell_prob_mn, E_q6_cell_prob_mn)

# q1_norm <- normalize(hydro_states, q1_cell_probs)
# q2_norm <- normalize(hydro_states, q2_cell_probs)
# q3_norm <- normalize(hydro_states, q3_cell_probs)
# q4_norm <- normalize(hydro_states, q4_cell_probs)
# q5_norm <- normalize(hydro_states, q5_cell_probs)
# q6_norm <- normalize(hydro_states, q6_cell_probs)

# q1_prob_mn_norm <- q1_cell_prob_mn * q1_norm
# q2_prob_mn_norm <- q2_cell_prob_mn * q2_norm
# q3_prob_mn_norm <- q3_cell_prob_mn * q3_norm
# q4_prob_mn_norm <- q4_cell_prob_mn * q4_norm
# q5_prob_mn_norm <- q5_cell_prob_mn * q5_norm
# q6_prob_mn_norm <- q6_cell_prob_mn * q6_norm

# row.names(q1_prob_mn_norm) <- row.names(q1_cell_prob_mn)
# row.names(q2_prob_mn_norm) <- row.names(q2_cell_prob_mn)
# row.names(q3_prob_mn_norm) <- row.names(q3_cell_prob_mn)
# row.names(q4_prob_mn_norm) <- row.names(q4_cell_prob_mn)
# row.names(q5_prob_mn_norm) <- row.names(q5_cell_prob_mn)
# row.names(q6_prob_mn_norm) <- row.names(q6_cell_prob_mn)

# if (exists("q6_prob_mn_norm")) q_all_mn_norm_e <- rbind(q1_prob_mn_norm, q2_prob_mn_norm, q3_prob_mn_norm, q4_prob_mn_norm, q5_prob_mn_norm, q6_prob_mn_norm)
# if (!exists("q6_prob_mn_norm")) q_all_mn_norm_e <- rbind(q1_prob_mn_norm, q2_prob_mn_norm, q3_prob_mn_norm, q4_prob_mn_norm, q5_prob_mn_norm)

#### Write csv file output ####
# write.csv(q1_prob_mn_norm, file="output/site3/q1_prob_mn_norm_e.csv")
# write.csv(q2_prob_mn_norm, file="output/site3/q2_prob_mn_norm_e.csv")
# write.csv(q3_prob_mn_norm, file="output/site3/q3_prob_mn_norm_e.csv")
# write.csv(q4_prob_mn_norm, file="output/site3/q4_prob_mn_norm_e.csv")
# write.csv(q5_prob_mn_norm, file="output/site3/q5_prob_mn_norm_e.csv")
# write.csv(q6_prob_mn_norm, file="output/site3/q6_prob_mn_norm_e.csv")

# Export File
write.csv(E_q_all_prob_mn, file="output/site3/E_q_all_prob_mn.csv")
E_q_all_prob_mn2 <- read.csv("output/site3/E_q_all_prob_mn.csv")
colnames(E_q_all_prob_mn2) <- c("cell","mean_prob")
E_q_all_prob_mn2 <- E_q_all_prob_mn2[order(E_q_all_prob_mn2$cell),]
write.csv(E_q_all_prob_mn2, file="output/site3/E_q_all_prob_mn.csv", row.names=FALSE)


####R Notebooks####
Execute code chunks by clicking the *Run* button within the chunk or by placing the cursor inside it and pressing *Cmd+Shift+Enter*. Add a new chunk by clicking the *Insert Chunk* button on the toolbar or by pressing *Cmd+Option+I*. Whenthe notebook is saved, an HTML file containing the code and output will be saved alongside it (click the *Preview* button or press *Cmd+Shift+K* to preview the HTML file).

####References ####
Gentleman R, Whalen E, Huber W and Falcon S (2017). graph: graph: A package to handle graph data structures. R package version 1.54.0.

Hansen KD, Gentry J, Long L, Gentleman R, Falcon S, Hahne F and Sarkar D (2017). Rgraphviz: Provides plotting capabilities for R graph objects. R package version 2.20.0.

Huber, W., Carey, J. V, Gentleman, R., Anders, S., Carlson, M., Carvalho, S. B, Bravo, C. H, Davis, S., Gatto, L., Girke, T., Gottardo, R., Hahne, F., Hansen, D. K, Irizarry, A. R, Lawrence, M., Love, I. M, MacDonald, J., Obenchain, V., Ole's, K. A, Pag'es, H., Reyes, A., Shannon, P., Smyth, K. G, Tenenbaum, D., Waldron, L., Morgan and M. (2015). “Orchestrating high-throughput genomic analysis with Bioconductor.” Nature Methods, 12(2), pp. 115–121. http://www.nature.com/nmeth/journal/v12/n2/full/nmeth.3252.html.

Koller D, Friedman N (2009). Probabilistic Graphical Models: Principles and Techniques. MIT Press.

Pearl J (2009). Causality: Models, Reasoning and Inference. Cambridge University Press, 2nd edition.

Nagarajan, Radhakrishnan, Marco Scutari, and Sophie Lèbre. Bayesian Networks in R. New York, NY: Springer New York, 2013. http://link.springer.com/10.1007/978-1-4614-6446-4. 
http://www.bnlearn.com

Scutari, Marco (2018). bnlearn: Bayesian Network Structure Learning, Parameter Learning and Inference. R package version 4.4 https://cran.r-project.org/web/packages/bnlearn/

https://stackoverflow.com/questions/40364458/r-some-conditional-probability-distributions-of-node-do-not-sum-to-one

example - https://stackoverflow.com/questions/34127885/bnlearn-wrong-dimensions-for-node

cpquery - http://www.bnlearn.com/documentation/man/cpquery.html

soft/hard evidence - https://pdfs.semanticscholar.org/13f5/795c9d4075d4468131fe9405bae5ec3e3857.pdf

